---
title: "608_JR_FInal_Project"
author: "Jeyaraman Ramalingam"
date: "5/21/2021"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
    toc_float: yes
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(maps)
library(dplyr)
library(ggplot2)
library(Amelia)
library(kableExtra)
```

## Overview

Terrorism is a never ending nemesis for the Global Community. Every Year , the threat because of terrorism is growing making it essential to look back and find out the vulnerable regions, countries and provinces/states. This would provide the government organizations necessary insights to take better decisions on handling the Global Terrorism.

## Data Source

Data Source URL - https://gtd.terrorismdata.com/files/gtd-1970-2019-4/

The GTD is an open-source database, which provides information on domestic and international terrorist attacks around the world since 1970, and now includes more than 200,000 events. For each event, a wide range of information is available, including the date and location of the incident, the weapons used, nature of the target, the number of casualties, and – when identifiable – the group or individual responsible.

```{r,message=FALSE,warning=FALSE,echo=F}
data<- read.csv("globalterrorismdb_csv.csv", header=TRUE, check.names = FALSE)
colnames(data)
```

The Input datasource has lot of unnecessary columns which are not required for this research. Hence the next step is to drop the columns apart from the below columns

* iyear
* imonth
* iday
* country
* country_txt
* region
* region_txt
* provstate
* city
* attacktype1
* attacktype1_txt
* attacktype2
* attacktype2_txt
* attacktype3
* attacktype3_txt

## Data Exploration and Preparation

### Data Subsetting

```{r,message=FALSE,warning=FALSE,echo=F}
df <- data %>%
  select(iyear,imonth,iday,country,country_txt,region,region_txt,provstate,city,attacktype1,attacktype1_txt,attacktype2,attacktype2_txt,attacktype3,attacktype3_txt)

df %>% head() %>% kable()
df_sample <- df %>% head(50)

```

### Summary Statistics

```{r,message=FALSE,warning=FALSE,echo=F}
summary(df)
```


### Missing Data Check

```{r,message=FALSE,warning=FALSE,echo=F}
colSums(is.na(df))
missmap(df, main="Missing Values")
```

The miss map output and Column Summaries shows us that the attack type 2 and 3 fields are not significant for the research. The next step is to drop the insignificant columns from the dataset

### Data Subsetting - Missing Columns

```{r,message=FALSE,warning=FALSE,echo=F}
df_final <- df %>% select(iyear,imonth,iday,country,country_txt,region,region_txt,provstate,city,attacktype1,attacktype1_txt)
colSums(is.na(df_final))
```

The Column Summaries shows us that there are no missing values. 


## Data Visualization

These Plots helps us to get some insights on the research questions. We can look at these visuals and compare against the Interactive Shiny App results. The Plots are generated based on the sample data

The below Plots are generated for the Terrorism Data Understanding.

### Terrorist Events by Country

The total number of terrorist events happened for each Country. This helps us identifying which countries are impacted.

```{r,message=FALSE,warning=FALSE,echo=F}
ggplot(df_sample, aes(x = country_txt)) + geom_dotplot(method="histodot", binwidth = 0.2,stackratio=0.8,stroke=2,fill="red")
```

### Terrorist Events by Region

The total number of terrorist events happened for each Region. This helps us identifying which region are impacted.

```{r,message=FALSE,warning=FALSE,echo=F}
ggplot(df_sample, aes(x = region_txt)) + geom_dotplot(method="histodot", binwidth = 0.2,stackratio=0.8,stroke=2,fill="green")
```

### Terrorist Events by AttackType

The total number of terrorist events happened for each attack type. This helps us identifying different types of attack used by terrorists.

```{r,message=FALSE,warning=FALSE,echo=F}
ggplot(df_sample, aes(x = attacktype1_txt)) + geom_dotplot(method="histodot", binwidth = 0.2,stackratio=0.8,stroke=2,fill="blue")
```


### Yearly Trend of Terrorist Events

The trend report of Year vs terrorist events. By looking at this plot we can clearly understand that the terrorism grew after 2010 and slowly reducing in 2020s.

```{r,message=FALSE,warning=FALSE,echo=F}
df_country_cnt <- df_final %>%
  group_by(country_txt) %>%
  summarise(event_cnt=n())

df_year_cnt <- df_final %>%
  group_by(iyear) %>%
  summarise(event_cnt=n())

ggplot(df_year_cnt, aes(x=iyear, y=event_cnt)) +  geom_point(aes(size=event_cnt),shape=18,color="red")  +   geom_smooth(method=lm,  linetype="dashed",  color="darkgreen", fill="gray") + xlab("Year") + ylab("Water Resource Scale")   + ggtitle("Yearly trend of Water Resource impact over Child Mortality")
```

Also from the Plot there is a huge spike in terrorist events after 2010. Next step is to identify which countries are impacted post 2010.

### Top Countries Impacted Post 2010

```{r,message=FALSE,warning=FALSE,echo=F}
df_country_2010 <- df_final %>%
  filter(iyear > 2010) %>%
  group_by(country_txt) %>%
  summarise(event_cnt=n()) %>%
  arrange(desc(event_cnt)) %>%
  head(10)

ggplot(df_country_2010, aes(x="", y=event_cnt, fill=country_txt)) +
geom_bar(stat="identity", width=1, color="black") +
coord_polar("y", start=0) +
theme_void() +
scale_fill_brewer(palette="Spectral")
```

Looking at the Plot The terrorism in Iraq is more after 2010 eventhough India ranks top in terrorism events

### Global Impact of Terrorism

This plot provides a heat map of World based on the terrorist events. India is the country which is severely impacted by terrorist events other than any country.

```{r,message=FALSE,warning=FALSE,echo=F}
df_country_cnt$region <- df_country_cnt$country_txt

df_country_cnt <- df_country_cnt %>%
  mutate(region = ifelse(region == "United States", "USA", region)) %>%
  mutate(region = ifelse(region == "Russian Federation", "Russia", region)) %>%
  mutate(region = ifelse(region == "Iran (Islamic Republic of)", "Iran", region)) %>%
  mutate(region = ifelse(region == "Kazakstan", "Kazakhstan", region))

world_map <- map_data("world")

terror_map <- left_join(world_map, df_country_cnt, by = "region")

ggplot(terror_map, aes(map_id = region, fill = event_cnt))+
  geom_map(map = terror_map,  color = "white")+
  expand_limits(x = terror_map$long, y = terror_map$lat)+
  scale_fill_gradient(low='#edaf98', high='#fa0511')

```


### Shiny Application

## Question 1

```{r,message=FALSE,warning=FALSE,echo=F,eval=FALSE}
ui <- fluidPage(

  
  dateRangeInput("daterange4", "Pick Date Range ",
                 start = df$event_date,
                 end = df$event_date),
  actionButton("do", "Fetch Report"),
  mainPanel(
    plotOutput("country_by_date"),width = 500  
    
  )
)

server <- function(input, output) {
  
  observeEvent(input$do, {
    output$country_by_date <- renderPlot({
      df_final %>%
        filter(event_date >= input$daterange4[1] & event_date < input$daterange4[2]) %>%
        select(country_txt) %>%
        group_by(country_txt) %>%
        summarise(event_cnt=n()) %>%
        arrange(desc(event_cnt)) %>%
        head(10) %>%
		ggplot(aes(country_txt,event_cnt,size=event_cnt,color=factor(event_cnt))) + geom_point() +   coord_flip()      
    })
  })
}

shinyApp(ui, server)
```

## Question 2

```{r,message=FALSE,warning=FALSE,echo=F,eval=FALSE}
ui <- fluidPage(
  
  selectInput("variables", "Variable:", choices=c("All",unique(df_country_list$country_txt))),
  dateRangeInput("daterange4", "Pick Date Range ",
                 start = df$event_date,
                 end = df$event_date),
  actionButton("do", "Fetch Report"),
  
  mainPanel(
    plotOutput("country_by_date"),width = 500  
    
  )
)

server <- function(input, output) {
  
  observeEvent(input$do, {
    output$country_by_date <- renderPlot({
      df_final %>%
        filter(event_date >= input$daterange4[1] & event_date < input$daterange4[2] & country_txt==input$variables) %>%
        select(provstate) %>%
        group_by(provstate) %>%
        summarise(event_cnt=n()) %>%
        arrange(desc(event_cnt)) %>%
        head(10) %>%
		ggplot(aes(provstate,event_cnt,size=event_cnt,color=factor(event_cnt))) + geom_col() +   coord_flip() 
      
    })
  })
}

shinyApp(ui, server)
```

## Question 3

```{r,message=FALSE,warning=FALSE,echo=F,eval=FALSE}
ui <- fluidPage(

  dateRangeInput("daterange4", "Pick Date Range ",
                 start = df$event_date,
                 end = df$event_date),
  actionButton("do", "Fetch Report"),
  
  mainPanel(
    plotOutput("country_by_date"),width = 500  
    
  )
)

server <- function(input, output) {
  
  observeEvent(input$do, {
    output$country_by_date <- renderPlot({
      df_final %>%
        filter(event_date >= input$daterange4[1] & event_date < input$daterange4[2]) %>%
        select(region_txt) %>%
        group_by(region_txt) %>%
        summarise(event_cnt=n()) %>%
        arrange(desc(event_cnt)) %>%
        head(10) %>%
        ggplot(aes(region_txt,event_cnt)) + geom_col(color="red",fill="gray",size=1)
      
    })
  })
}

shinyApp(ui, server)
```

## Question 4

```{r,message=FALSE,warning=FALSE,echo=F,eval=FALSE}
ui <- fluidPage(
  
  selectInput("variables", "Variable:", choices=c("All",unique(df_final$attacktype1_txt))),
  
  dateRangeInput("daterange4", "Pick Date Range ",
                 start = df$event_date,
                 end = df$event_date),
  actionButton("do", "Fetch Report"),
  mainPanel(
    plotOutput("country_by_date"),width = 500  
    
  )
)

server <- function(input, output) {
  
  observeEvent(input$do, {
    output$country_by_date <- renderPlot({
      df_final %>%
        filter(event_date >= input$daterange4[1] & event_date < input$daterange4[2] & attacktype1_txt==input$variables) %>%
		  select(country_txt) %>%
	  mutate(region=country_txt) %>%
	  group_by(region) %>%
	  summarise(event_cnt=n()) %>%
	  arrange(desc(event_cnt)) %>%
	  head(10) %>%
	  right_join(world_map, by = "region") %>%
	  ggplot(aes(map_id = region, fill = event_cnt))+  geom_map(map = terror_map,  color = "white")+  expand_limits(x = terror_map$long, y = terror_map$lat)+  scale_fill_gradient(low='#edaf98', high='#fa0511')
	   })
  })
}

shinyApp(ui, server)
```